### 2021-10-26 ml/Ergasia1.Rmd
ML: Ergasia1: Added Ask1 with testthat unit tests, which were added to renv.lock

% !TeX program = XeLaTeX
---
title: "Εργασία 1"
author: "Θωμάς Παππάς"
date: "26/10/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::read_chunk('R/ergasia1.R')
```

# Άσκηση 1
```{r ask1_nnk}
```