### 2024-12-09 thesis/slides.tex
Thesis: Pres: First entry with template, agenda and citations

\documentclass{beamer}
\usepackage[style=authoryear,backend=bibtex]{biblatex}
\addbibresource{ref.bib}

\mode<presentation>
{
	\usetheme[progressbar=foot,numbering=fraction,background=light,block=fill]{metropolis}
	\usecolortheme{default}
	\usefonttheme{default}
	\setbeamertemplate{navigation symbols}{}
	\setbeamertemplate{caption}[numbered]
	\setbeamertemplate{frame footer}{Thomas Pappas | ALMA }
}

\title[]{Pricing Games in Heterogeneous Parallel Networks}
\author[Thomas Pappas]{Thomas Pappas}
\institute[ALMA]{
	\begin{columns}
		\column{0.4\textwidth}
		\begin{figure}
			\centering
			\includegraphics[scale=0.1]{alma.png}
		\end{figure}
		\column{0.5\textwidth}
		ALMA \\
		\tiny{INTER-INSTITUTIONAL GRADUATE PROGRAM \\
			"ALGORITHMS, LOGIC AND DISCRETE MATHEMATICS"}
	\end{columns}
}
\date{\today}

\begin{document}
\AtBeginSection[]{
  \begin{frame}[noframenumbering,plain]
	  \vfill
	  \centering
	  \begin{beamercolorbox}[sep=8pt,center,shadow=true,rounded=true]{title}
	    \usebeamerfont{title}\insertsectionhead\par
	  \end{beamercolorbox}
	  \vfill
  \end{frame}
}

\begin{frame}[noframenumbering,plain]
  \titlepage
\end{frame}

\begin{frame}[noframenumbering,plain]
    \frametitle{Agenda}
    \tableofcontents[hideallsubsections]
\end{frame}

\section{History}

\begin{frame}{History}
	\begin{itemize}
		\item Selfish routing \footcite{pigou1920economics}
		\item Congestion games \footcite{Rosenthal1973ACO}
		\item Wardrop equilibrium \footcite{wardrop_theoretical_1952}
		\item Pricing competition \footcite{10.1287/moor.1060.0231}
		\begin{itemize}
			\item Homogeneous case with toll caps \footcite{Harks_2019}
		\end{itemize}
	\end{itemize}
\end{frame}

\section{Homogeneous Pricing Games}

\section{Sensitivity split}

\section{Pseudo-heterogeneous Pricing Games}

\section{Step distribution function}

\section{Future work}

\begin{frame}[noframenumbering,plain]{}
	\centering
    \huge Thank you!\\
    \normalsize Thomas Pappas\\
    thpappas@di.uoa.gr
\end{frame}

\end{document}